---
title: "Homework #1 Solutions"
author: "Doni Obidov"
date: "2023-09-04"
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: false # Remove the "#" before section numbers
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load the data and the packages
```{r}
data(chickwts) # Load the chickwts dataset
library(ggplot2) # Load ggplot2
```

# 1
```{r}
t.conf.int <- function(x, conf.level = 0.95) {
  n <- length(x) # Sample size
  alpha <- 1 - conf.level # Confidence level
  df <- n - 1 # Degrees of freedom
  t_score <- qt(1 - alpha / 2, df) # t-score
  
  mean <- mean(x) # Sample mean
  std <- sd(x) # Sample standard deviation
  err.margin <- t_score * (std / sqrt(n)) # Margin of error
  
  lwr.CI <- mean - err.margin  # Lower bound of the confidence interval
  upr.CI <- mean + err.margin  # Upper bound of the confidence interval
  
  # Return vector
  result <- c(
    lwr.CI = lwr.CI,
    upr.CI = upr.CI,
    mean = mean,
    err.margin = err.margin,
    conf.level = conf.level
  )
  
  return(result)
}
```

```{r}
# Test the function:
set.seed(5761)
data <- rnorm(30, mean = 0, sd = 1)
confidence_interval <- t.conf.int(data, conf.level = 0.85)
print(confidence_interval)
```

# 2

## a
```{r}
# Create a list of chick weights by feed type
chick_weights_by_feed <- split(chickwts$weight, chickwts$feed)

chick_weights_by_feed
```

## b
```{r}
horsebean_weights <- chick_weights_by_feed$horsebean

# Calculate the 95% confidence interval
confidence_interval <- t.conf.int(horsebean_weights)
confidence_interval
```

## c
```{r}
# Calculate the adjusted confidence level
alpha_adjusted <- 0.05 / 6 # Bonferroni correction
conf_level_adjusted <- 1 - alpha_adjusted

# Calculate simultaneous confidence intervals using lapply
confidence_intervals_lapply <- lapply(chick_weights_by_feed, function(weights) {
  t.conf.int(weights, conf.level = conf_level_adjusted)
})

# Print the confidence intervals
print(confidence_intervals_lapply)

# Calculate simultaneous confidence intervals using sapply
confidence_intervals_sapply <- sapply(chick_weights_by_feed, function(weights) {
  t.conf.int(weights, conf.level = conf_level_adjusted)
})

# Print the confidence intervals
print(confidence_intervals_sapply)
```

The main difference between `lapply` and `sapply` is in the structure of the output:

- `lapply` returns a list where each element corresponds to a different feed type, and each element contains a vector with the lower bound, upper bound, mean, margin of error, and confidence level for that feed type.

- `sapply` returns a matrix where each column represents a different feed type, and each row corresponds to the lower bound, upper bound, mean, margin of error, and confidence level.

## d
```{r}
# Create the ggplot object
p <- ggplot(chickwts, aes(x = feed, y = weight, fill = feed)) +
  geom_boxplot() + # Create boxplots
  labs(title = "Chick Weights by Feed Type", x = "Feed Type", y = "Weight (Gram)") + # Add title and axis labels
  theme_minimal() + # Choose a plot theme
  theme(plot.title = element_text(hjust = 0.5, size = 16))

print(p)
```

## e
```{r}
# Create the ggplot object
p <- ggplot(chickwts, aes(x = feed, y = weight, fill = feed, color = feed)) +
  geom_violin(alpha = 0.7, width = 0.5) + # Create a violin plot
  geom_jitter(alpha = 0.7, width = 0.15) + # Add jittered points
  labs(title = "Chick Weights by Feed Type", x = "Feed Type", y = "Weight (Gram)") +  # Add title and axis labels
  theme_minimal() + # Choose a plot theme
  theme(plot.title = element_text(hjust = 0.5, size = 16))

# Display the plot
print(p)
```

## f
Boxplots provide a quick summary of the data's central tendency and spread. This makes it easier to compare different feed types. Moreover, boxplots clearly indicate outlines.
Violin plots, on the other hand, provide a more detailed view of data distribution, letting us explore the shape and density of the data. In particular, violin lots offer insights into the shape and multimodality of the data.