---
title: 'MA5732-GLMs Module 4 Homework #2'
author: "Doni Obidov"
output:
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    theme: journal
    toc: yes
---

```{r}
# Load the packages
library(GLMsData)
library(statmod)
library(MASS)
library(ggplot2)
library(tidyverse)

# Load the dataset
data(germ)
data(germBin)
```

# Problem 4.7

## Exercise 1

```{r}
# Create a dataframe
data <- tibble(
  Substance = c(rep("A", 4), rep("B", 5)),
  Dose = c(1.0, 20.0, 100.0, 200.0, 1.0, 62.5, 125.0, 250.0, 500.0),
  No_cell_Samples = c(400, 200, 200, 200, 400, 200, 200, 200, 200),
  No_Cells_Aberrant = c(3, 5, 14, 4, 5, 2, 2, 4, 7)
)

data$Substance <- factor(data$Substance)
```

```{r}
# Fit a binomial GLM
model <- glm(cbind(No_Cells_Aberrant, No_cell_Samples - No_Cells_Aberrant) ~ Substance + Dose, data = data, family = binomial(link = "logit"))

# Summarize the model
summary(model)
```

```{r}
# Test statistic for Substance B
z_statistic <- coef(summary(model))["SubstanceB", "z value"]

# Extract p-value for Substance B
p_value <- coef(summary(model))["SubstanceB", "Pr(>|z|)"]

# Print
cat("Wald Z test:\n")
cat("Z test statistic:", z_statistic, "\n")
cat("P-value for Substance B:", p_value, "\n")
```

**Conclusion: there is a significant difference between substance a and substance b.**

## Exercise 2

```{r}
# Fit a binomial GLM
model_1 <- glm(cbind(No_Cells_Aberrant, No_cell_Samples - No_Cells_Aberrant) ~ Substance + log(Dose), data = data, family = binomial(link = "logit"))

# Summarize the model
summary(model_1)
```

**log(Dose) model: AIC = 49.123, residual deviance: 13.329**

**Dose model: AIC = 52.448, residual deviance: 16.653**

**Based on both AIC and residual deviance, the log(Dose) model is better**

## Exercise 3

```{r}
# Extract coefficient for log(Dose)
coefficient_log_dose <- coef(summary(model_1))["log(Dose)", "Estimate"]

# Extract standard error for log(Dose)
se_log_dose <- coef(summary(model_1))["log(Dose)", "Std. Error"]

# Calculate 95% confidence interval for log(Dose)
lower_ci_log_dose <- coefficient_log_dose - 1.96 * se_log_dose
upper_ci_log_dose <- coefficient_log_dose + 1.96 * se_log_dose

# Print coefficient and its 95% confidence interval
cat("Coefficient for log(Dose):", coefficient_log_dose, "\n")
cat("95% Confidence Interval for log(Dose): [", lower_ci_log_dose, ",", upper_ci_log_dose, "]\n")
```

```{r}
# Calculate odds ratio for dose increase from 20 to 200
odds_ratio <- exp(coefficient_log_dose * (log(200) - log(20)))

# Print odds ratio
cat("If dose increases from 20 to 200, odds increase by a factor :", odds_ratio, "\n")
```

## Exercise 4

**The presence of multiple substances complicates the dose-response relationship, making it difficult to interpret the ED50 accurately.**

# Problem 4.8

## Exercise 1

```{r}
# Fit logistic regression model using binomial GLM
binomial_model <- glm(Germ/Total ~ Seeds + Extract, data = germ, family = binomial(link = "logit"), weights = Total)
# binomial_model <- glm(cbind(Germ, Total-Germ) ~ Seeds + Extract, data = germ, family = binomial(link = "logit"))

# Fit logistic regression model using Bernoulli GLM
bernoulli_model <- glm(factor(Result, levels = c("NotGerm", "Germ")) ~ Seeds + Extract, data = germBin, family = binomial(link = "logit"))
```

```{r}
# Parameter estimates and standard errors for binomial GLM
cat("Parameter Estimates and Standard Errors for Binomial GLM:\n")
print(summary(binomial_model)$coefficients)

# Parameter estimates and standard errors for Bernoulli GLM
cat("\nParameter Estimates and Standard Errors for Bernoulli GLM:\n")
print(summary(bernoulli_model)$coefficients)
```

**Parameter estimates and standard errors from two models are the same.**

## Exercise 2

```{r}
# Extracting null and residual deviances from the summary of binomial model
binomial_null_deviance <- summary(binomial_model)$null.deviance
binomial_residual_deviance <- summary(binomial_model)$deviance

# Display the results
cat("Residual Deviance, Null Deviance, and Difference for Binomial GLM:\n")
cat("Residual Deviance:", binomial_residual_deviance, "\n")
cat("Null Deviance:", binomial_null_deviance, "\n")
cat("Difference:", binomial_null_deviance - binomial_residual_deviance, "\n\n")

# Extracting null and residual deviances from the summary of bernoulli model
bernoulli_null_deviance <- summary(bernoulli_model)$null.deviance
bernoulli_residual_deviance <- summary(bernoulli_model)$deviance

# Display the results
cat("Residual Deviance, Null Deviance, and Difference for Bernoulli GLM:\n")
cat("Residual Deviance:", bernoulli_residual_deviance, "\n")
cat("Null Deviance:", bernoulli_null_deviance, "\n")
cat("Difference:", bernoulli_null_deviance - bernoulli_residual_deviance, "\n")
```

**Although the residual deviances and the null deviances are different, their differences are the same.**

## Exercise 3

```{r}
# Perform sequential likelihood-ratio test for binomial model
binomial_anova <- anova(binomial_model, test = "Chisq", dispersion = 1)

# Print results for binomial model
cat("Sequential Likelihood-Ratio Test for Binomial GLM:\n")
print(binomial_anova)

# Perform sequential likelihood-ratio test for bernoulli model
bernoulli_anova <- anova(bernoulli_model, test = "Chisq", dispersion = 1)

# Print results for bernoulli model
cat("\nSequential Likelihood-Ratio Test for Bernoulli GLM:\n")
print(bernoulli_anova)
```

**Although the residual deviances and the associated df are different, the p-values are the same.**

## Exercise 5 (bonus)

```{r}
# Log-likelihood for binomial model
binomial_log_likelihood <- logLik(binomial_model)

# Log-likelihood for Bernoulli model
bernoulli_log_likelihood <- logLik(bernoulli_model)

# Print log-likelihoods
cat("Log-Likelihood for Binomial Model:", binomial_log_likelihood, "\n")
cat("Log-Likelihood for Bernoulli Model:", bernoulli_log_likelihood, "\n")

# Difference of log-likelihoods
log_likelihood_difference <- binomial_log_likelihood - bernoulli_log_likelihood

# Print the difference
cat("Difference of Log-Likelihoods (Binomial - Bernoulli):", log_likelihood_difference, "\n")
```

```{r}
theoretical <- function(Germ, Total) {
  sum_result <- 0
  for (i in 1:length(Germ)) {
    mi <- Total[i]
    yi <- Germ[i] / Total[i]
    sum_result <- sum_result + log(choose(mi, mi * yi))
  }
  return(sum_result)
}

# Calculate the sum
theoretical_dif <- theoretical(germ$Germ, germ$Total)

# Print the result
cat("Theoretical difference:", theoretical_dif, "\n")
```