---
title: 'MA5732-GLMs Homework #2'
author: "Doni Obidov"
output:
  html_document:
    toc: yes
    df_print: paged
  html_notebook:
    theme: journal
    toc: yes
---

```{r}
# Load the packages
# install.packages("GLMsData")
library(GLMsData)
library(ggplot2)

# Load the datasets
data(crawl)
```

# Problem 1.3

## Exercise 1

```{r}
# Scatter plot of Age against Temp (Data Aggregated over Month)
ggplot(crawl, aes(x = Temp, y = Age)) +
  geom_point() +
  labs(x = "The Monthly Average Temperature (°F) Six Months After the Birth Month",
       y = "The Mean Age (Weeks) at Which the Babies Started to Crawl",
       title = "Scatter Plot of Age against Temperature (Data Aggregated over Month)") +
  theme_minimal()
```

With an exception of a few outliers, the mean age to crawl is linearly reducing with an increasing monthly average temperature. The homoscedasticity does not seem to be a huge issue, if the outliers are ignored.

## Exercise 2

The response variable **Age** is the **mean age** (in completed weeks) at which the babies born this month started to crawl, **not ages of individual babies**. If ages of individual babies at age $x_i$ have variance $𝜎^2$, then mean ages of babies from the same group have variance $𝜎^2/w_i$, where $w_i$ is the **sample size** of group $i$.

## Exercise 3

A possible model for the data is: 
\[y_i \sim N(\mu_i, \frac{\sigma^2}{w_i}) = N(\beta_0 + \beta_1 x_{i1}, \frac{\sigma^2}{w_i})\]
where $(i=1,⋯,n)$ and $y_i$ are independent.

```{r}
# Fit a weighted linear regression model
weighted_lm <- lm(Age ~ Temp, data = crawl, weights = 1/SampleSize)

# Display the summary of the model
summary(weighted_lm)
```

```{r}
# Store the results in variables
coefficient_estimate <- coef(weighted_lm)["Temp"]
standard_error <- summary(weighted_lm)$coefficients["Temp", "Std. Error"]
t_value <- coef(weighted_lm)["Temp"] / summary(weighted_lm)$coefficients["Temp", "Std. Error"]
p_value <- coef(summary(weighted_lm))["Temp", "Pr(>|t|)"]
```

### Interpretation:

- The estimated coefficient for `Temp` (`r round(coefficient_estimate, 5)`) represents the expected change in the mean age at which babies start to crawl for each one-degree Fahrenheit increase in the monthly average temperature six months after birth.

- The standard error (`r round(standard_error, 5)`) provides a measure of uncertainty or variability in this estimate. It indicates how much the estimated coefficient might vary if the study were repeated.

- The t-value (`r round(t_value, 3)`) is the ratio of the estimated coefficient to its standard error. In this case, it suggests that the coefficient for `Temp` is significantly different from zero.

- The p-value (`r round(p_value, 4)`) associated with `Temp` is less than the conventional significance level of `0.05`, indicating that there is evidence to reject the null hypothesis that the coefficient for `Temp` is zero. In other words, there is evidence to suggest that the temperature has a significant effect on the mean age at which babies start to crawl.

In summary, the model suggests that there is a significant negative relationship between the monthly average temperature and the mean age at which babies start to crawl. As the temperature increases, the mean age to crawl tends to decrease.

## Exercise 4

```{r}
# Degrees of freedom (df) for the t-distribution
df <- length(crawl$Temp) - 2 # n-(p+1)

# Calculate the t-test statistic
t_statistic <- (coefficient_estimate - 0) / standard_error

# Calculate the one-sided p-value for a left-tailed test
p_value_one_sided <- pt(t_statistic, df)

# Display the results
t_statistic
p_value
```

```{r}
# Check if we reject the null hypothesis based on the one-sided p-value
if (p_value_one_sided < 0.05) {
  cat("The one-sided p-value is less than 0.05. Reject the null hypothesis.\n")
  cat("Conclusion: There is evidence to suggest that babies take longer to learn to crawl in colder months.")
} else {
  cat("The one-sided p-value is greater than or equal to 0.05. Fail to reject the null hypothesis.\n")
  cat("Conclusion: There is insufficient evidence to suggest that babies take longer to learn to crawl in colder months.")
}
```

## Exercise 5

```{r}
# Given information
alpha = 0.05  # Significance level
alpha_over_2 = alpha / 2  # Two-tailed alpha

# Critical value from t-distribution
critical_value = qt(1 - alpha_over_2, df)

# Calculate the margin of error
margin_of_error = critical_value * standard_error

# Calculate the lower and upper bounds of the confidence interval
lower_bound = coefficient_estimate - margin_of_error
upper_bound = coefficient_estimate + margin_of_error

# Display the results
cat("95% Confidence Interval for the Slope:\n")
cat(sprintf("Lower Bound: %.5f\n", lower_bound))
cat(sprintf("Upper Bound: %.5f\n", upper_bound))
```

**95% Confidence Interval for the Slope:**

The 95% confidence interval for the slope of the relationship between monthly average temperature (`Temp`) and the mean age at which babies start to crawl is approximately \([-0.13574, -0.02303]\).

**Interpretation:**

1. **Negative Slope:** The interval is entirely below zero, indicating a negative slope for the relationship. This suggests that as the monthly average temperature increases, the mean age at which babies start to crawl tends to decrease.

2. **Statistical Significance:** Since the interval does not include zero, we can interpret this result as statistically significant. There is evidence at the 95% confidence level to suggest that there is a significant negative relationship between monthly average temperature and the mean age at which babies start to crawl.

In summary, based on the 95% confidence interval, we can conclude that there is a statistically significant negative association between temperature and the mean age at which babies start to crawl. As temperatures increase, the mean age at which babies begin to crawl tends to decrease.

## Exercise 6

```{r}
# Fit unweighted linear regression model
unweighted_lm <- lm(Age ~ Temp, data = crawl)

# Scatter plot with both weighted and unweighted regression lines
ggplot(crawl, aes(x = Temp, y = Age, color = factor(c('Weighted', 'Unweighted')))) + 
  geom_point(color = "black") +
  geom_smooth(aes(linetype = "Weighted"), method = "lm", formula = y ~ x, se = FALSE, color = "blue", linetype = "solid", linewidth = 1) +
  geom_smooth(aes(linetype = "Unweighted"), method = "lm", formula = y ~ x, se = FALSE, color = "red", linetype = "dashed", linewidth = 1) +
  labs(x = "The Monthly Average Temperature (°F) Six Months After the Birth Month",
       y = "The Mean Age (Weeks) at Which the Babies Started to Crawl",
       title = "Comparison of Weighted (Blue) and Unweighted (Red) Regression Lines") + theme_minimal() +
theme(legend.position = "bottom")
```

There is almost no difference between the weighted and not weighted linear models in this case. This can be possibly be explained by a relatively small difference in sample sizes between different groups.

# Problem 1.4

```{r}
# Given parameter estimates and standard errors
constant_estimate <- 100.812
constant_se <- 13.096

age_estimate <- 0.332
age_se <- 0.062

waist_estimate <- 0.411
waist_se <- 0.090

alcohol_estimate <- -3.003
alcohol_se <- 1.758

smoking_estimate <- -0.362
smoking_se <- 2.732

temp_estimate <- -0.521
temp_se <- 0.262
```

## Exercise 1

```{r}
# Compute p-values for Age and Smoking
# Null hypothesis: βj = 0 vs Alternative hypothesis: βj ≠ 0
df <- 574 - 6
p_value_age <- 2 * (pt(abs(age_estimate / age_se), df = df, lower.tail = FALSE))
p_value_smoking <- 2 * (pt(abs(smoking_estimate / smoking_se), df = df, lower.tail = FALSE))

# Display p-values
cat("P-value for Age:", p_value_age, "\n")
cat("P-value for Smoking:", p_value_smoking, "\n")

# Comment on the p-values
if (p_value_age < 0.05) {
  cat("P-value for Age is less than 0.05. Reject the null hypothesis.\n")
  cat("Conclusion: Age has a significant effect on systolic blood pressure.\n")
} else {
  cat("P-value for Age is greater than or equal to 0.05. Fail to reject the null hypothesis.\n")
  cat("Conclusion: There is insufficient evidence to suggest that Age has a significant effect on systolic blood pressure.\n")
}

if (p_value_smoking < 0.05) {
  cat("P-value for Smoking is less than 0.05. Reject the null hypothesis.\n")
  cat("Conclusion: Smoking has a significant effect on systolic blood pressure.\n")
} else {
  cat("P-value for Smoking is greater than or equal to 0.05. Fail to reject the null hypothesis.\n")
  cat("Conclusion: There is insufficient evidence to suggest that Smoking has a significant effect on systolic blood pressure.\n")
}
```

## Exercise 2

After adjusting for Age, Waist circumference, Alcohol consumption, and Smoking habits the relationship between Ambient temperature and systolic blood pressure can be interpreted as follows:

For each **one-degree Celsius increase in Ambient temperature**, **on average**, the systolic **blood pressure** is expected to **decrease by 0.521** mm Hg, holding other variables constant.

## Exercise 3

```{r}
# Compute a 95% confidence interval for the regression parameter for Ambient temperature
alpha = 0.05  # Significance level
alpha_over_2 = alpha / 2  # Two-tailed alpha

# Degrees of freedom for t-distribution
df_temp = 574 - 6

# Critical value from t-distribution
critical_value_temp = qt(1 - alpha_over_2, df_temp)

# Calculate the margin of error
margin_of_error_temp = critical_value_temp * temp_se

# Calculate the lower and upper bounds of the confidence interval
lower_bound_temp = temp_estimate - margin_of_error_temp
upper_bound_temp = temp_estimate + margin_of_error_temp

# Display the results
cat("95% Confidence Interval for the regression parameter for Ambient temperature:\n")
cat(sprintf("Lower Bound: %.5f\n", lower_bound_temp))
cat(sprintf("Upper Bound: %.5f\n", upper_bound_temp))
```

The 95% confidence interval for the regression parameter for Ambient temperature is \([-1.03561, -0.00639]\). Since this interval does not include zero, we can interpret it as follows:

- **Negative Interval:** The entire interval is below zero, indicating a negative effect of Ambient temperature on systolic blood pressure.

- **Statistical Significance:** Since the interval does not contain zero, we can conclude that there is statistical evidence at the 95% confidence level to suggest that Ambient temperature has a significant effect on systolic blood pressure in Ghanaian men aged between 18 and 65.

In summary, based on the confidence interval, we can infer that each one-degree Celsius increase in Ambient temperature is associated with a statistically significant decrease in systolic blood pressure, after adjusting for other variables in the model.

## Exercise 4

```{r}
# Predict the mean systolic blood pressure for 35-year-old Ghanaian men
# (who do not smoke, do drink alcohol, and have a waist circumference of 100 cm)
# when the ambient temperature is 30°C

# Given values
age_pred <- 35
waist_pred <- 100
alcohol_pred <- 1  # Yes: 1
smoking_pred <- 0  # No: 0
temp_pred <- 30

# Calculate the predicted mean systolic blood pressure
predicted_pressure <- constant_estimate +
  age_estimate * age_pred +
  waist_estimate * waist_pred +
  alcohol_estimate * alcohol_pred +
  smoking_estimate * smoking_pred +
  temp_estimate * temp_pred

cat(sprintf("Predicted mean systolic blood pressure for 35-year-old Ghanaian men:\n%.2f mm Hg\n", predicted_pressure))
```